Structured Nonlinear Feedback Design with Separable Control-Contraction Metrics

### 2016-07-25 R0V0.tex
> Removed the Reviews from Commit.

{\itshape Stabilizability Notion.} A forward complete solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be  \emph{globally exponentially uniformly stabilizable} with \emph{rate} $\lambda>0$ if there exist a constant value $C>0$ and a feedback law $k^\ast:\mathbb{R}_{\geq0}\times\mathbb{R}^n\to\mathbb{R}^m$, denoted as $k^\ast(\cdot,\cdot,X^\ast,u^\ast)$, such that the inequality
\begin{equation}\label{eq:global exponential uniform stabilizability}
	\left|X(t,x^\ast,u^\ast)-X(t,x,k^\ast)\right|\leq Ce^{-\lambda t}|x^\ast-x|
\end{equation}
holds, for every $t\geq0$, and for every $x\in\mathbb{R}^n$ (\cite{Manchester2014a}).

@@

\begin{definition}\label{def:US}
	The system \eqref{eq:general system} is said to be \emph{universally stabilizable} with rate $\lambda$ if, for every forward complete solution $X^\ast(\cdot,\cdot,u^\ast)$ to \eqref{eq:general system}, there exists a static feedback law $k^\ast$ for system \eqref{eq:global exponential uniform stabilizability} that globally exponentially uniformly stabilizes it.
\end{definition}

Note that Definition \ref{def:US} reduces to the notion of stabilizability of equilibria, when $x^\ast=0$ (for further reading on stabilizability, the reader may address \cite{Bacciotti:1992}).

@@

The property described in item 2 of Problem \ref{problem formulation} is particularly relevant for the design feedback laws with a prescribed structure (topology). Consider the network composed of systems described by the following equation.
\begin{equation}\label{eq:example}
	\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i.
	\end{array}\right.
\end{equation}
As explained in Section \ref{sec:Illustration}, when more than five of these systems are connected, the CCM approach cannot compute a unconstrained controller on a standard desktop machine, due to the number of variables. However, by exploring sparsity, up to 512 of theses systems can be connected and the controller can be computed.

@@

\begin{definition}\label{def:}
		 A smooth function $V:\mathbb{R}^n\times\mathbb{R}^n\to\mathbb{R}_{\geq0}$ is said to be a \emph{metric for system \eqref{eq:general system}} if there exist constant values $\underline{c}>$ and $\overline{c}>$ such that the inequality
		 \begin{subequations}\label{eq:metric}
		 \begin{equation}
		 	\underline{c}|\delta_x|^2\leq V(x,\delta_x)\leq \overline{c}|\delta_x|^2,
	 	 \end{equation}
	 	 holds, for every $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$. Given fixed controls $u$ and $\delta u$ for systems \eqref{eq:general system} and \eqref{eq:general system:differential}, respectively. A metric system \eqref{eq:general system} receives the adjective \emph{contraction} if there exists a value $\lambda>0$ such that the inequality
	 	 \begin{equation}
	 	 	\frac{dV}{dt}(X(t,x,u),\Delta(t,x,\delta x,u,\delta u))\leq -\lambda V(x,\delta_x)
 	 	 \end{equation}
 	 	 \end{subequations}
 	 	 holds, for every pair $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$.
\end{definition}
Note that a bounded Riemannian metric defined, for every $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$, as $V(x,\delta_x)=|\delta_x|_x^2$ and satisfying the set of inequalities \eqref{eq:metric} is a contraction metric for system \eqref{eq:general system}. With an abuse of concept, from now the bounded Riemannian metric defined above will be called simply as \emph{metric}.

The existence a contraction metric for system \eqref{eq:general system} with $u\equiv0$ implies that every two solutions of this system converge to each other exponentially. The proof of this claim can be found in \cite[Theorem 1]{Lewis1951}, and \cite[Theorems 5.7 and 5.33]{Reich2005}, and \cite[Lemma 3.3]{Isac2008}.

@@

\begin{definition}[{\cite{Manchester2014a}}]\label{def:}
	A metric for system \eqref{eq:general system} is said to be a \emph{control-contraction metric for system \eqref{eq:general system}} if there exists a constant value $\lambda>0$ such that the condition
	\begin{subequations}\label{eq:Arstein-Sontag}
		\begin{equation}\label{eq:}
			\delta_x^\top M(x)B(x)=0
		\end{equation}
		implies that the inequality
		\begin{align}
			\delta_x^\top(\dot{M}+A^\top M+MA)\delta_x\leq\ -2\lambda \delta_x^\top M\delta_x\label{eq:}
		\end{align}
		holds, where $\dot{M}:=\partial_{f+Bu}M$.
	\end{subequations}
\end{definition}
The set of equations \eqref{eq:Arstein-Sontag} is an adaptation of Artstein-Sontag's condition for exponential convergence of solutions. Given a control-contraction metric for system \eqref{eq:general system}, Finsler's lemma (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) provides stabilizing a feedback law of the form $\delta_u=K\delta_x$ for system \eqref{eq:general system:differential} defined for every $\delta_x\in\mathbb{R}^n$, where $K:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{n\times m}$.

@@

Step 1 (Offline MI computation). Consider the change of variables $\eta=M\delta$ and define the matrix $W=M^{-1}$. The set of equations \eqref{eq:Arstein-Sontag} is equivalent (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) to the existence of a bounded differentiable function $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ such that $W=W^\top\succ0$ and a function $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ satisfying the following matrix inequality (MI)
\begin{equation}\label{eq:MI formulation}
	-\dot{W}+AW+WA^\top+BY+(BY)^\top+2\lambda W\preceq0,
\end{equation}
that is linear on the matrix variables $W$ and $Y$, for every $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$. Consequently, $M=W^{-1}$ is a control-contraction metric for system \eqref{eq:general system}.

@@

\begin{corollary}\label{cor:distributed computation}
	 If there exist a smooth functions $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ such that, for every $(x,\delta x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$, $W=W^\top\succ0$, $W$ is block-diagonal, $Y\in\Xi_Y$ and the pair $(W,Y)$ is a solution to the MI
	\begin{equation}\label{eq:MI formulation:i}
		T_{ii}\preceq 0
	\end{equation}
	for every index $i\in\mathbb{N}_{[1,N]}$, and the matrix $T$ is diagonally dominant, then pair of matrices $W$ and $Y$ is a solution to the MI \eqref{eq:MI formulation}.
\end{corollary}

@@

When $N\geq6$ and $Y$ is a full matrix with elements described by polynomials, the MI \eqref{eq:MI formulation} could not be solved with the optimization parser Yalmip \cite{Loefberg2004,Loefberg2009} and the solver Mosek running on an Intel Core i7, 8GB RAM, Microsoft Windows 10 and on Matlab 2015b. A reason for this issue is the dependence of each element of the matrix $Y$ on the variable $q$. Another reason is the need to verify if the matrix \eqref{eq:MI formulation} is negative definite. When no structural constraints are imposed on $Y$, this verification can lead $n^2$ equations to be solved. This motivates the approach proposed in this paper.

@@

	For every $\delta_\eta\in\mathbb{R}^n$, consider the product $\delta_\eta^\top T\delta_\eta$. Each line $i\in\mathbb{N}_{[1,N]}$ of this product is given by
	\begin{equation*}
		\delta_{\eta_i}^\top\begin{bmatrix}
			T_{i1}&\cdots&T_{i(i-1)}&T_{ii}&T_{i(i+1)}&\cdots&T_{iN}
		\end{bmatrix}\delta_\eta.
	\end{equation*}
	Inequalities \eqref{eq:BDD2} and \eqref{eq:MI formulation:i} imply that the inequalities
	\begin{equation*}
		\sum_{\substack{j=1\\ j\neq i}}^N\delta_{\eta_i}^\top T_{ij}\delta_{x_j}\leq \delta_{\eta_i}^\top T_{ii}\delta_{\eta_i}\leq 0
	\end{equation*}
	hold, for every $\delta_\eta\in\mathbb{R}^n$. Consequently, the inequality $\delta_\eta^\top T\delta_\eta\leq 0$ holds, for every $\delta_\eta\in\mathbb{R}^n$. This later inequality is equivalent to the MI \eqref{eq:MI formulation}. This concludes the proof of Corollary~\ref{cor:distributed computation}.
\end{proof}

>>>>

{\itshape Stabilizability Notion.} A forward complete solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be  \emph{globally exponentially uniformly stabilizable} with \emph{rate} $\lambda>0$ if there exist a constant value $C>0$ and a feedback law $k^\ast:\mathbb{R}_{\geq0}\times\mathbb{R}^n\to\mathbb{R}^m$, denoted as $k^\ast(\cdot,\cdot,X^\ast,u^\ast)$, such that the inequality
\begin{equation}\label{eq:global exponential uniform stabilizability}
	\left|X(t,x^\ast,u^\ast)-X(t,x,k^\ast)\right|\leq Ce^{-\lambda t}|x^\ast-x|
\end{equation}
holds, for every $t\geq0$, and for every $x\in\mathbb{R}^n$ (\cite{Manchester2014a}). Equivalently, if inequality \eqref{eq:global exponential uniform stabilizability} hold, then the solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be \emph{globally exponentially uniformly stable} for system \eqref{eq:general system} in closed loop with $k^\ast$

@@

\begin{definition}\label{def:US}
	The system \eqref{eq:general system} is said to be \emph{universally stabilizable} with rate $\lambda$ if, for every forward complete solution $X^\ast(\cdot,\cdot,u^\ast)$ to \eqref{eq:general system}, there exists a static feedback law $k^\ast$ for system \eqref{eq:global exponential uniform stabilizability} rendering $X^\ast(\cdot,\cdot,u^\ast)$ globally exponentially uniformly stable with rate $\lambda$.
\end{definition}

@@

The property described in item 2 of Problem \ref{problem formulation} is particularly relevant for the design feedback laws with a prescribed structure (topology). Consider the network composed of systems described by the following equation.
\begin{equation}\label{eq:example}
	\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i.
	\end{array}\right.
\end{equation}
As explained in Section \ref{sec:Illustration}, when more than five of these systems are connected, the CCM approach was unable to provide unconstrained controller on a standard desktop machine, due to the number of variables. However, by exploring sparsity, up to 512 of theses systems can be connected and the controller can be computed.

@@

\begin{definition}\label{def:}
		 A smooth function $V:\mathbb{R}^n\times\mathbb{R}^n\to\mathbb{R}_{\geq0}$ is said to be a \emph{metric for system \eqref{eq:general system}} if there exist constant values $\underline{c}>$ and $\overline{c}>$ such that the inequality
		 \begin{subequations}\label{eq:metric}
		 \begin{equation}
		 	\underline{c}|\delta_x|^2\leq V(x,\delta_x)\leq \overline{c}|\delta_x|^2,
	 	 \end{equation}
	 	 holds, for every $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$. Given fixed controls $u$ and $\delta u$ for systems \eqref{eq:general system} and \eqref{eq:general system:differential}, respectively. A metric system \eqref{eq:general system} receives the adjective \emph{contraction} if there exists a value $\lambda>0$ such that the inequality
	 	 \begin{equation}
	 	 	\frac{dV}{dt}(X(t,x,u),\Delta(t,x,\delta x,u,\delta u))\leq -\lambda V(x,\delta_x)
 	 	 \end{equation}
 	 	 \end{subequations}
 	 	 holds, for every pair $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$.
\end{definition}
Note that a bounded Riemannian metric defined, for every $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$, as $V(x,\delta_x)=|\delta_x|_x^2$ and satisfying the set of inequalities \eqref{eq:metric} is a contraction metric for system \eqref{eq:general system}. For brevity, from now the bounded Riemannian metric defined above will be called simply as \emph{metric}.

The existence a contraction metric for system \eqref{eq:general system} implies that every two solutions of this system converge to each other exponentially. The proof of this claim, for the case in which $u\equiv0$, can be found in \cite[Theorem 1]{Lewis1951}, and \cite[Theorems 5.7 and 5.33]{Reich2005}, and \cite[Lemma 3.3]{Isac2008}.

@@

\begin{definition}[{\cite{Manchester2014a}}]\label{def:}
	A metric for system \eqref{eq:general system} is said to be a \emph{control-contraction metric for system \eqref{eq:general system}} if there exists a constant value $\lambda>0$ such that the condition
	\begin{subequations}\label{eq:Arstein-Sontag}
		\begin{equation}\label{eq:}
			\delta_x\neq0,\quad\delta_x^\top M(x)B(x)=0
		\end{equation}
		implies that the inequality
		\begin{align}
			\delta_x^\top(\dot{M}+A^\top M+MA)\delta_x<\ -2\lambda \delta_x^\top M\delta_x\label{eq:}
		\end{align}
		holds, where $\dot{M}:=\partial_{f+Bu}M$.
	\end{subequations}
\end{definition}
The set of equations \eqref{eq:Arstein-Sontag} is an adaptation of Artstein-Sontag's condition for exponential convergence of solutions. Given a control-contraction metric for system \eqref{eq:general system}, Finsler's lemma (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) provides stabilizing a feedback law of the form $\delta_u=K\delta_x$ for system \eqref{eq:general system:differential} defined for every $\delta_x\in\mathbb{R}^n$, where $K:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{n\times m}$.

@@

Step 1 (Offline MI computation). Consider the change of variables $\eta=M\delta$ and define the matrix $W=M^{-1}$. The set of equations \eqref{eq:Arstein-Sontag} is equivalent (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) to the existence of a bounded differentiable function $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ such that $W=W^\top\succ0$ and a function $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ satisfying the following matrix inequality (MI)
\begin{equation}\label{eq:MI formulation}
	-\dot{W}+AW+WA^\top+BY+(BY)^\top+2\lambda W\prec0,
\end{equation}
that is linear on the matrix variables $W$ and $Y$, for every $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$. Consequently, $M=W^{-1}$ is a control-contraction metric for system \eqref{eq:general system}.

@@

\begin{corollary}\label{cor:distributed computation}
	 If there exist a smooth functions $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ such that, for every $(x,\delta x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$, $W=W^\top\succ0$, $W$ is block-diagonal, $Y\in\Xi_Y$ and the pair $(W,Y)$ is a solution to the MI
	\begin{equation}\label{eq:MI formulation:i}
		T_{ii}\prec 0
	\end{equation}
	for every index $i\in\mathbb{N}_{[1,N]}$, and the matrix $T$ is diagonally dominant, then pair of matrices $W$ and $Y$ is a solution to the MI \eqref{eq:MI formulation}.
\end{corollary}

@@

When $N\geq6$ and $Y$ is a full matrix with elements described by polynomials, the MI \eqref{eq:MI formulation} could not be solved with the optimization parser Yalmip \cite{Loefberg2004,Loefberg2009} and the solver Mosek running on an Intel Core i7, 8GB RAM, Microsoft Windows 10 and on Matlab 2015b. A reason for this issue is the dependence of each element of the matrix $Y$ on the variable $q$. Another reason is the need to verify if the matrix \eqref{eq:MI formulation} is negative definite. When no structural constraints are imposed on $Y$, this verification can lead to
\begin{equation*}
\left(\sum_{i=1}^N\begin{pmatrix}
n_i+d\\ d
\end{pmatrix}\right)^2
\end{equation*}
equations to be solved. This motivates the approach proposed in this paper.

@@

	For every $\delta_\eta\in\mathbb{R}^n$, consider the product $\delta_\eta^\top T\delta_\eta$. Each line $i\in\mathbb{N}_{[1,N]}$ of this product is given by
	\begin{equation*}
		\delta_{\eta_i}^\top\begin{bmatrix}
			T_{i1}&\cdots&T_{i(i-1)}&T_{ii}&T_{i(i+1)}&\cdots&T_{iN}
		\end{bmatrix}\delta_\eta.
	\end{equation*}
	Inequalities \eqref{eq:BDD2} and \eqref{eq:MI formulation:i} imply that the inequalities
	\begin{equation*}
		\sum_{\substack{j=1\\ j\neq i}}^N\delta_{\eta_i}^\top T_{ij}\delta_{x_j}\leq \delta_{\eta_i}^\top T_{ii}\delta_{\eta_i}< 0
	\end{equation*}
	hold, for every $\delta_\eta\in\mathbb{R}^n$. Consequently, the inequality $\delta_\eta^\top T\delta_\eta\leq 0$ holds, for every $\delta_\eta\in\mathbb{R}^n$. This later inequality is equivalent to the MI \eqref{eq:MI formulation}. This concludes the proof of Corollary~\ref{cor:distributed computation}.
\end{proof}
